---
title: "FW_438"
Author: rose
Details: looking at the first 438 turkers' responses on the fw turk survey
date: CSV downloaded on 9/29/14
---

#Some preliminaries - libraries, data, etc. 
```{r}
library(ggplot2)
library(plyr)
library(reshape2)
theme_set(theme_bw())
df=read.csv("FW_590_data.csv")
```


ANALYSIS STARTS HERE
-----

#Freqs of first words - count - these are the standardized words!
```{r}
freqs <- ddply(df, .(first_wd_standard), summarise, count=length(first_wd_standard))
freqs$first_wd_standard <- factor(freqs$first_wd_standard, levels=with(freqs, first_wd_standard[order(count, first_wd_standard, decreasing = TRUE)]))
q <- qplot(first_wd_standard, count, geom="bar", stat="identity", 
      data=subset(freqs, count>5))
q + theme(axis.text.x = element_text(angle=90, hjust = 1, vjust=.5))
```

#Props of first words - standardized!
```{r}
freqs$prop <- freqs$count/sum(freqs$count)
q <- qplot(first_wd_standard, prop, geom="bar", stat="identity", 
           data=subset(freqs, count>5))
q + theme(axis.text.x=element_text(angle=90, hjust=1, vjust=.5))+ylab("Proportion of Total Words")
```

#For comparison, here are the original first words - proportions
```{r}
freqs <- ddply (df,  .(first_wd_original), summarise, count=length(first_wd_original))

freqs$first_wd_original <- factor(freqs$first_wd_original, 
                     levels=unique(with(freqs, 
                                 first_wd_original[order(count, 
                                            first_wd_original, decreasing = TRUE)])))
freqs$prop <- freqs$count/sum(freqs$count)
q <- qplot(first_wd_original, prop, data=subset(freqs, count>5), geom="bar", position="dodge", stat="identity")
q + theme(axis.text.x=element_text(angle=90, hjust=1, vjust=.5))+ylab("Proportion of Total Words")
```

#Proportions of Ages 

```{r}
freqs <- ddply(df, .(word_age), summarise, count=length(word_age))
freqs$word_age <- factor(freqs$word_age, levels=with(freqs, 
          word_age[order(count, word_age, decreasing = TRUE)]))
freqs$prop <- freqs$count/sum(freqs$count)
q <- qplot(word_age, prop, data=freqs, geom="bar", position="dodge", stat="identity")
q + theme(axis.text.x=element_text(angle=90, hjust=1, vjust=.5))+ylab("Proportion of Ages")
```

```{r}
library(stringr)
df$age <- as.numeric(str_replace(as.character(df$word_age)," months",""))
qplot(age,breaks=0:25,
      data=subset(df,first_wd_standard != "N/A"))
df$age.grp <- cut(df$age,breaks=c(6,10,14,18))
```



#Proportions of Word~Age
```{r}
freqs <- ddply(df, .(first_wd_standard, word_age), summarise, count=length(first_wd_standard))
freqs <- ddply(freqs, .(first_wd_standard), mutate, total.count=sum(count))
freqs <- ddply(freqs, .(word_age), mutate, prop=count/sum(count))
freqs$prop <- freqs$count/sum(freqs$count)

freqs$first_wd_standard <- factor(freqs$first_wd_standard, 
                                  levels=unique(with(freqs,
                                  first_wd_standard[order(total.count,  first_wd_standard, decreasing=TRUE)])))

q <- qplot(first_wd_standard, prop, fill=word_age, position="dodge",
           geom="bar", stat="identity", data=subset(freqs, total.count>5
                                                    & word_age != "don't remember"))
q + theme(axis.text.x=element_text(angle=90, hjust=1, vjust=.5))

```

CDI category proportions 

```{r}
freqs <- ddply(df, .(word_type), summarise, 
               count=length(word_type))
freqs$prop <- freqs$count/sum(freqs$count)

qplot(word_type, prop, geom="bar", stat="identity", 
           data=subset(freqs, count>1)) + 
  theme(axis.text.x=element_text(angle=90, hjust=1, vjust=.5))+
  ylab("Proportion of Total Words")

```

By age group:

```{r}
freqs <- ddply(df, .(word_type,age.grp), summarise, 
               count=length(word_type))
freqs$word_type <- factor(freqs$word_type, 
                     levels=unique(with(freqs, 
                                 word_type[order(count, 
                                            word_type, decreasing = TRUE)])))
freqs <- ddply(freqs, .(age.grp), summarise,
               word_type = word_type,
               prop=count/sum(count))

qplot(word_type, prop, geom="bar", stat="identity", 
           data=subset(freqs, prop>.01)) + 
  theme(axis.text.x=element_text(angle=90, hjust=1, vjust=.5))+
  ylab("Proportion of Total Words")+ 
  facet_wrap(~age.grp)
```

#Words by gender 
```{r}
freqs <- ddply(df, .(first_wd_standard, gender), summarise, count=length(first_wd_standard))
freqs <- ddply(freqs, .(first_wd_standard), mutate, total.count=sum(count))
freqs$first_wd_standard <- factor(freqs$first_word_standard, 
                                  levels=unique(with(freqs, first_wd_standard
                                                     [order(total.count,
                                                            first_wd_standard,
                                                            decreasing = TRUE)])))
freqs$prop <- freqs$count/sum(freqs$count)
q <- qplot(first_wd_standard, prop, fill=gender, 
           position="dodge",
           geom="bar",
           stat="identity",
           data=subset(freqs, total.count>5 & gender!=""))
q + theme(axis.text.x = element_text(angle=90, hjust = 1, vjust=.5)) + 
  ylab("Proportion of Children")
```

#Looking within gender - females
```{r}
freqs <- ddply(df, .(first_wd_standard, gender), summarise, count=length(first_wd_standard))
freqs <- ddply(freqs, .(gender), mutate, total.count=sum(df$gender=="female"))
freqs$females <- freqs$gender == "female"
freqs$first_wd_standard <- factor(freqs$first_word_standard, 
                                  levels=unique(with(freqs, first_wd_standard
                                                     [order(total.count,
                                                            first_wd_standard,
                                                            decreasing = TRUE)])))
freqs$prop <- freqs$count/(freqs$total.count)
q <- qplot(first_wd_standard, prop, fill=females, 
           position="dodge", 
           geom="bar",
           stat="identity", 
           data=subset(freqs, count>5 & gender!="male"))
q + theme(axis.text.x = element_text(angle=90, hjust = 1, vjust=.5)) + 
  ylab("Proportion of Females")
```

#Looking within gender - males
```{r}
freqs <- ddply(df, .(first_wd_standard, gender), summarise, count=length(first_wd_standard))
freqs <- ddply(freqs, .(gender), mutate, total.count=sum(df$gender=="male"))
freqs$males <- freqs$gender == "male" 
freqs$first_wd_standard <- factor(freqs$first_wd_standard, levels=unique(with(freqs,
                                        first_wd_standard[order(total.count, first_wd_standard,
                                                                decreasing=TRUE)])))
freqs$prop <- freqs$count/(freqs$total.count)
q <- qplot(first_wd_standard, prop, fill=males, 
           position="dodge", 
           geom="bar",
           stat="identity", 
           data=subset(freqs, count>5 & gender!="female"))
q + theme(axis.text.x = element_text(angle=90, hjust = 1, vjust=.5)) + 
  ylab("Proportion of Males")
```

#Males and females together 
```{r}
freqs <- ddply(df, .(first_wd_standard, gender), summarise, count=length(first_wd_standard))
freqs <- ddply(freqs, .(gender), mutate, total.count=sum(length(df$gender)))
freqs$prop <- freqs$count/(freqs$total.count)
q <- qplot(first_wd_standard, prop, fill=gender, 
           position="dodge", 
           geom="bar",
           stat="identity", 
           data=subset(freqs, count>5))
q + theme(axis.text.x = element_text(angle=90, hjust = 1, vjust=.5)) + 
  ylab("Proportion of Total Children")
```

#Grouping the Ages - I'm grouping them in groups of 3 (except for 0-3, which is 4) for right now just for ease of analysis. However, I think we should really look at this data and try to figure out a less arbitrary way of grouping things. 
```{r}
df$group1 <- df$word_age == "0 months" | df$word_age == "1 month" | df$word_age == "2 months" | df$word_age == "3 months"
df$group2 <- df$word_age == "4 months" | df$word_age == "5 months" | df$word_age == "6 months"
df$group3 <- df$word_age == "7 months" | df$word_age == "8 months" | df$word_age == "9 months"
df$group4 <- df$word_age == "10 months" | df$word_age == "11 months" | df$word_age == "12 months"
df$group5 <- df$word_age == "13 months" | df$word_age == "14 months" | df$word_age == "15 months"
df$group6 <- df$word_age == "16 months" | df$word_age == "17 months" | df$word_age == "18 months"
df$group7 <- df$word_age == "19 months" | df$word_age == "20 months" | df$word_age == "21 months"
df$group8 <- df$word_age == "22 months" | df$word_age == "23 month" | df$word_age == "24+ months"


```

#Age split, but not really...not sure how to do that yet...but this is 7-9m 
```{r}
freqs <- ddply(df, .(first_wd_standard, group3), summarise, count=length(first_wd_standard))
freqs <- ddply(freqs, .(first_wd_standard), mutate, total.count=sum(count))
freqs <- ddply(freqs, .(group3), prop=count/sum(count))
freqs$first_wd_standard <- factor(freqs$first_wd_standard, levels=unique(with(freqs,                                                                          first_wd_standard[order(total.count, first_wd_standard,                                                                                                      decreasing=TRUE)])))
freqs$prop <- freqs$count/sum(freqs$count)
q <- qplot(first_wd_standard, prop, fill = group3, 
           position="dodge", geom="bar", stat="identity", 
           data=subset(freqs, total.count>5))
q + theme(axis.text.x = element_text(angle=90, hjust = 1, vjust=.5))

```

#Same, but this is for group 4 - 10-12m
```{r}
freqs <- ddply(df, .(first_wd_standard, group4), summarise, count=length(first_wd_standard))
freqs <- ddply(freqs, .(first_wd_standard), mutate, total.count=sum(count))
freqs <- ddply(freqs, .(group4), prop=count/sum(count))
freqs$first_wd_standard <- factor(freqs$first_wd_standard, levels=unique(with(freqs,                                                                          first_wd_standard[order(total.count, first_wd_standard,                                                                                                      decreasing=TRUE)])))
freqs$prop <- freqs$count/sum(freqs$count)
q <- qplot(first_wd_standard, prop, fill = group4, 
           position="dodge", geom="bar", stat="identity", 
           data=subset(freqs, total.count>5))
q + theme(axis.text.x = element_text(angle=90, hjust = 1, vjust=.5))

```

#same, 13-15m
```{r}
freqs <- ddply(df, .(first_wd_standard, group5), summarise, count=length(first_wd_standard))
freqs <- ddply(freqs, .(first_wd_standard), mutate, total.count=sum(count))
freqs <- ddply(freqs, .(group5), prop=count/sum(count))
freqs$first_wd_standard <- factor(freqs$first_wd_standard, levels=unique(with(freqs,                                                                          first_wd_standard[order(total.count, first_wd_standard,                                                                                                      decreasing=TRUE)])))
freqs$prop <- freqs$count/sum(freqs$count)
q <- qplot(first_wd_standard, prop, fill = group5, 
           position="dodge", geom="bar", stat="identity", 
           data=subset(freqs, total.count>5))
q + theme(axis.text.x = element_text(angle=90, hjust = 1, vjust=.5))

```

```{r}


```
